---
title: "from book 'ggplot2 essential' movie examples simplified"
author: "Илья Кочергин"
date: '29 апреля 2016 г '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library}
library(ggplot2)
library(ggplot2movies) # Набор данных movies вынесен из новых версий ggplot2
library(dplyr)
```
## Наборы данных, используемые в  примерах
## Movie dataset (пакет ggplot2)

Набор данных movies переместили в отдельный пакет `ggplot2movies` из пакета `ggplot2` в последних версиях.

#### Описание из пакета `ggplot2movies`:

A data frame with 28819 rows and 24 variables

* title. Title of the movie.

* year. Year of release.

* budget. Total budget (if known) in US dollars

* length. Length in minutes.

* rating. Average IMDB user rating.

* votes. Number of IMDB users who rated this movie.

* r1-10. Multiplying by ten gives percentile (to nearest 10%) of users who rated this movie a 1.

* mpaa. MPAA rating.

* action, animation, comedy, drama, documentary, romance, short. Binary variables representing
if movie was classified as belonging to that genre.

```{r show movies data stucture}
glimpse(movies)

```

```{r reshape movies dataset}
d1 <-data.frame(movies[movies$Action==1, c("budget", "Short",
"year")])
d1$Type <- "Animation"
d2 <-data.frame(movies[movies$Animation==1, c("budget", "Short",
"year")])
d2$Type <- "Animation"
d3 <-data.frame(movies[movies$Comedy==1, c("budget", "Short",
"year")])
d3$Type <- "Comedy"
d4 <-data.frame(movies[movies$Drama==1, c("budget", "Short",
"year")])
d4$Type <- "Drama"
d5 <-data.frame(movies[movies$Documentary==1, c("budget", "Short",
"year")])
d5$Type <- "Documentary"
d6 <-data.frame(movies[movies$Romance==1, c("budget", "Short",
"year")])
d6$Type <- "Romance"
myMovieData <- rbind(d1, d2, d3, d4, d5, d6)
names(myMovieData) <- c("Budget", "Short", "Year", "Type" )
```

Реструктуризованный набор данных (по жанрам)

```{r show myMovieData dataset stucture}
glimpse(myMovieData)
```

## Bar plots

```{r barplot1}
ggplot(myMovieData,aes(Type,fill=Type)) + geom_bar()
```

### Порядок столбцов

Для номинальных данных прядок расположения на оси X можно задать при помощи изменения порядка уровней (levels) в столбце типа фактор.

Упорядочим уровни (factor levels) для получения нужного порядка столбцов (по убыванию количества фильмов)

```{r order levels}
# myMovieData_o_by_cnt <- mutate(myMovieData,Type ,)
newLevelsMovieType <- sort(table(myMovieData$Type),decreasing = T)
newLevelsMovieType 
myMovieData_o_by_cnt <-
  mutate(myMovieData,
        Type=factor(Type,levels = names(newLevelsMovieType)))
ggplot(myMovieData_o_by_cnt,aes(Type,fill=Type)) + geom_bar()
```

### Раположение прямоугольников, соответствующих одному значению X

3 варианта `position_adjustment` для `geom_bar(position="...")` :

* stack -- один над другим

* dodge -- один рядом с другим 

* fill  -- заполняют всю высоту 

```{r bar position}
gbar <-  ggplot(data=myMovieData_o_by_cnt,
                aes(x=Type, fill=factor(Short)))
gbar + geom_bar(position="stack") # default 
gbar + geom_bar(position="dodge")
gbar + geom_bar(position="fill")
```

#### Пример с Boxplot 

```{r boxplot1}
gBoxPoint <- 
ggplot()
```

### Шкалы и статистики 

```{r remove NA}
myMovieData <- na.omit(myMovieData)
glimpse(myMovieData)
```

Сгруппируем по типу и подсчитаем средний бюджет

```{r group by Type}
myMovieByTypeByShort <- 
myMovieData %>% 
  group_by(Type,Short) %>% 
  summarise(meanBudget=mean(Budget),sumBudget=sum(as.numeric(Budget)))

ggBudgetsByType <- 
  ggplot(myMovieByTypeByShort,aes(x=Type,y=sumBudget,fill=factor(Short))) +
  geom_bar(stat="identity",position="dodge")
ggBudgetsByType
ggBudgetsByType +scale_y_log10()
```


**Замечание 1** `geom_bar()` неправильно воспринимается с логарифмической шкалой по Y.

**Замечание 2** `geom_bar(binwidth=1)` с ныне устаревшим параметром использовался для построения столбцового графика по качественной переменной на оси X. Теперь для количественной шкалы X используется `geom_histogram()`

```{r boxplot}
ggplot(data=myMovieData,
aes(Type,Budget)) +
geom_jitter() + geom_boxplot(alpha=I(0.6)) +
scale_y_log10()
```



#### Цветовые шкалы


выдержка из `help(package="RColorBrewer")`:

ColorBrewer is Copyright (c) 2002 **Cynthia Brewer**, Mark Harrower, and The Pennsylvania State University. All rights reserved. 
The ColorBrewer palettes have been included in this R package with permission of the copyright holder. 
`

There are 3 types of palettes, **sequential**, **diverging**, and **qualitative**.

1. *Sequential palettes** are suited to **ordered data** that progress from low to high. **Lightness steps** dominate the look of these schemes, with **light colors** for **low** data values to **dark colors** for **high** data **values**. 
2. *Diverging palettes* put equal emphasis on mid-range critical values and extremes at both ends of the data range. The critical class or **break in the middle** of the legend is emphasized with **light colors** and **low and high extremes** are emphasized with **dark colors** that have **contrasting hues**. 
3. *Qualitative palettes* do not imply magnitude differences between legend classes, and hues are used to create the primary visual differences between classes. Qualitative schemes are best suited to representing nominal or categorical data.

```{r display brewer palettes}
 library(RColorBrewer)
 display.brewer.all(type = "seq")   ## количественные переменные -- правый край шкалы тёмный
 display.brewer.all(type = "div")   ## количественные переменные -- края шкалы темные 
 display.brewer.all(type = "qual")  ## качественные переменные -- контрастные цвета

```




#### Параметр *alpha* -- Непрозрачность

```{r area plot}
movieByYearByType <- myMovieData %>%  group_by(Year,Type) %>% 
  summarise(meanBudget=mean(Budget)) %>% 
  arrange(Type,Year)
ggarea1 <- filter(movieByYearByType,Type %in% c('Drama','Comedy','Animation')) %>% 
  ggplot(data=., aes(x=Year,y=meanBudget,fill=Type)) + geom_area()
ggarea1
ggarea1+ scale_y_log10()
ggarea1+geom_area(alpha=0.5) + scale_y_log10()
```


